Modelling multi-scale, state-switching functional data with hidden Markov models: Supplement B (simulation study)

### 2021-10-26 Report/Supp_Sim.tex
Update Final Version

% Title, authors, affiliations
\title{Modelling state-switching functional data with hidden Markov models: supplementary material (simulation study)}%\query{Q1}
\date{}
\author{BlindedA, BlindedB}%\authorref{1}\thanksref{*}}
%\author{BlindedB}%\authorref{2}}
%\affiliation[1]{Author affiliations will go here in the accepted manuscript, but do NOT include them in your initial submission because it must be anonymous.}
%\affiliation[2]{Second Affiliation}

@@

    	\noindent Figure \arabic{fignum}: Acceleration data for a five dives of a selected simulated data set. The raw accelerometer data is denoted as $Y^*_{t,\tilde t^*}$ and recorded at 50 times per second while $\Zone_{t,\tilde t^*}$ and $\Ztwo_{t,\tilde t^*}$ are recorded at once every two seconds. Dives are separated by vertical black lines. The colour of the curve corresponds to the true subdive state, while the colour of the background corresponds to the true dive type. Subdive state 1 can plausibly be interpreted as ``gliding", subdive state 2 can plausibly be interpreted as active swimming, and subdive state 3 can be interpretted as sudden or aggressive movement.
    	\addtocounter{fignum}{1}

    \newpage
    \section{Accuracy comparisons}

@@

Model                       & \multicolumn{1}{c}{Train Time (m)} & \multicolumn{1}{c}{Dive Type} & \multicolumn{1}{c}{Subdive Type} & \multicolumn{1}{c}{Dive Accuracy} & \multicolumn{1}{c}{Subdive Accuracy}  \\ \hline

@@

        \noindent Table \arabic{tablenum}: Average decoding accuracies and training times for all models used to categorize dive type and subdive state in the simulation study. Each of the four models was fit to 500 training data sets comprised of 100 simulated dives and tested on test data sets also comprised of 100 simulated dives. Reported values are averages, and $\pm$ refers to the sample standard deviation across the 500 data sets. Rows labelled All/All correspond to overall average decoding accuracy.
        \addtocounter{tablenum}{1}

    \newpage
    \section{Empirical joint distributions - dive duration ($Y_t$)}

@@

Model                       & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Dive Type} & \multicolumn{1}{c}{Estimate}   & \multicolumn{1}{c}{Bias}   & \multicolumn{1}{c}{Empirical SE}   & \multicolumn{1}{c}{Observed Fischer SE}     \\ \hline
\multirow{4}{*}{CarHHMM-DFT}& \multirow{2}{*}{$\mu_Y$}      & 1                             & $27.33$                         & $0.07$ $(p=0.287)$          & $1.39$                             & $1.03 \pm 0.12$                             \\
                            &                               & 2                             & $127.14$                        & $0.07$ $(p=0.945)$          & $22.56$                             & $15.71 \pm 6.726$                             \\
                            & \multirow{2}{*}{$\sigma_Y$}   & 1                             & $10.80$                         & $-0.11$ $(p=0.035)$          & $1.15$                             & $0.84 \pm 0.11$                             \\
                            &                               & 2                             & $59.52$                         & $-4.54$ $(p=0.000)$          & $17.68$                             & $13.04 \pm 5.53$                             \\ \hline
\multirow{4}{*}{HHMM-DFT}   & \multirow{2}{*}{$\mu_Y$}      & 1                             & $27.37$                         & $0.10$ $(p=0.084)$          & $1.33$                             & $1.03 \pm 0.11$                             \\
                            &                               & 2                             & $126.84$                         & $-0.23$ $(p=0.818)$          & $22.00$                             & $15.37 \pm 6.11$                             \\
                            & \multirow{2}{*}{$\sigma_Y$}   & 1                             & $10.86$                         & $-0.05$ $(p=0.351)$          & $1.19$                             & $0.85 \pm 0.11$                             \\
                            &                               & 2                             & $57.14$                         & $-6.92$ $(p=0.000)$          & $16.12$                             & $12.62 \pm 5.17$                             \\ \hline
\multirow{4}{*}{CarHHMM}    & \multirow{2}{*}{$\mu_Y$}      & 1                             & $27.25$                         & $-0.02$ $(p=0.826)$          & $2.20$                             & $1.04 \pm 0.14$                             \\
                            &                               & 2                             & $117.62$                         & $-9.45$ $(p=0.000)$          & $34.34$                             & $15.13 \pm 7.10$                             \\
                            & \multirow{2}{*}{$\sigma_Y$}   & 1                             & $10.52$                         & $-0.39$ $(p=0.001)$          & $2.59$                             & $0.86 \pm 0.15$                             \\
                            &                               & 2                             & $56.98$                         & $-7.08$ $(p=0.000)$          & $16.33$                             & $12.53 \pm 5.95$                             \\ \hline

@@

        \noindent Table \arabic{tablenum}: Estimates and standard errors for the parameters of the distribution of dive duration across all four models in the simulation study. The ``Estimate" column is the average across 500 simulations, and the ``Observed Fisher SE" column is the median across 500 simulations. The $\pm$ refers to the inter-quartile range. SE stands for standard error. Reported $p$-values test if the observed bias is statistically significant using a one-sample $t$-test.
        \addtocounter{tablenum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHHMM-DFT and correspond to both dive types. The red star represents the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the HHMM-DFT and correspond to both dive types. The red star represents the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHHMM and correspond to both dive types. The red star represents the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of dive duration parameter estimates ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHMM-DFT. There is no red star which represents the true values because the CarHMM-DFT assumes that there is only one dive type.
        \addtocounter{fignum}{1}

    \newpage
    \section{Empirical joint distributions - acceleration ($\Zone_{t,\tilde t^*}$)}

@@

Model                       & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Subdive Type} & \multicolumn{1}{c}{Estimate}   & \multicolumn{1}{c}{Bias}   & \multicolumn{1}{c}{Empirical SE}   & \multicolumn{1}{c}{Observed Fischer SE}     \\ \hline
\multirow{9}{*}{CarHHMM-DFT}& \multirow{3}{*}{$\mu_A^*$}    & 1                                & $-0.012$                         & $-0.012$ $(p=0.004)$          & $0.093$                             & $0.066 \pm 0.038$                             \\
                            &                               & 2                                & $0.099$                         & $-0.001$ $(p=0.436)$          & $0.019$                             & $0.019 \pm 0.003$                             \\
                            &                               & 3                                & $0.200$                         & $0.000$ $(p=0.965)$          & $0.060$                             & $0.053 \pm 0.013$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.050$                         & $0.000$ $(p=0.376)$          & $0.002$                             & $0.002 \pm 0.000$                             \\
                            &                               & 2                                & $0.100$                         & $-0.000$ $(p=0.672)$          & $0.002$                             & $0.002 \pm 0.000$                             \\ 
                            &                               & 3                                & $0.298$                         & $-0.002$ $(p=0.024)$          & $0.016$                             & $0.015 \pm 0.002$                             \\
                            & \multirow{3}{*}{$\phi_A^*$}   & 1                                & $0.968$                         & $-0.002$ $(p=0.001)$          & $0.012$                             & $0.010 \pm 0.002$                             \\
                            &                               & 2                                & $0.828$                         & $-0.002$ $(p=0.007)$          & $0.017$                             & $0.017 \pm 0.002$                             \\
                            &                               & 3                                & $0.600$                         & $-0.010$ $(p=0.001)$          & $0.064$                             & $0.065 \pm 0.011$                             \\ \hline
\multirow{9}{*}{HHMM-DFT}   & \multirow{3}{*}{$\mu_A^*$}    & 1                                & $0.048$                         & $0.048$ $(p=0.000)$          & $0.033$                             & $0.007 \pm 0.001$                             \\
                            &                               & 2                                & $0.087$                         & $-0.013$ $(p=0.000)$          & $0.018$                             & $0.006 \pm 0.001$                             \\
                            &                               & 3                                & $0.215$                         & $0.015$ $(p=0.000)$          & $0.065$                             & $0.022 \pm 0.004$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.160$                         & $0.110$ $(p=0.000)$          & $0.023$                             & $0.005 \pm 0.001$                             \\
                            &                               & 2                                & $0.158$                         & $0.058$ $(p=0.000)$          & $0.010$                             & $0.004 \pm 0.000$                             \\ 
                            &                               & 3                                & $0.356$                         & $0.056$ $(p=0.000)$          & $0.035$                             & $0.015 \pm 0.002$                             \\

@@

\multirow{9}{*}{CarHHMM}    & \multirow{3}{*}{$\mu_A^*$}    & 1                                & $0.060$                         & $0.060$ $(p=0.000)$          & $0.034$                             & $0.019 \pm 0.006$                             \\
                            &                               & 2                                & $0.097$                         & $-0.003$ $(p=0.000)$          & $0.016$                             & $0.007 \pm 0.001$                             \\
                            &                               & 3                                & $0.201$                         & $0.001$ $(p=0.634)$          & $0.058$                             & $0.062 \pm 0.011$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.111$                         & $0.061$ $(p=0.000)$          & $0.004$                             & $0.003 \pm 0.000$                             \\
                            &                               & 2                                & $0.285$                         & $0.185$ $(p=0.000)$          & $0.009$                             & $0.005 \pm 0.000$                             \\ 
                            &                               & 3                                & $1.150$                         & $0.850$ $(p=0.000)$          & $0.080$                             & $0.048 \pm 0.009$                             \\
                            & \multirow{3}{*}{$\phi_A^*$}   & 1                                & $0.830$                         & $-0.140$ $(p=0.000)$          & $0.048$                             & $0.016 \pm 0.003$                             \\
                            &                               & 2                                & $0.004$                         & $-0.826$ $(p=0.000)$          & $0.012$                             & $0.001 \pm 0.023$                             \\
                            &                               & 3                                & $0.000$                         & $-0.610$ $(p=0.000)$          & $0.000$                             & $0.000 \pm 0.000$                             \\ \hline
\multirow{9}{*}{CarHMM-DFT} & \multirow{3}{*}{$\mu_A^*$}    & 1                                & $-0.008$                         & $-0.008$ $(p=0.142)$          & $0.114$                             & $0.064 \pm 0.037$                             \\
                            &                               & 2                                & $0.101$                         & $0.001$ $(p=0.438)$          & $0.019$                             & $0.019 \pm 0.003$                             \\
                            &                               & 3                                & $0.200$                         & $0.000$ $(p=0.926)$          & $0.057$                             & $0.054 \pm 0.014$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.050$                         & $-0.000$ $(p=0.005)$          & $0.001$                             & $0.002 \pm 0.000$                             \\
                            &                               & 2                                & $0.100$                         & $-0.000$ $(p=0.000)$          & $0.002$                             & $0.002 \pm 0.000$                             \\ 
                            &                               & 3                                & $0.299$                         & $-0.001$ $(p=0.117)$          & $0.015$                             & $0.015 \pm 0.002$                             \\
                            & \multirow{3}{*}{$\phi_A^*$}   & 1                                & $0.968$                         & $-0.002$ $(p=0.001)$          & $0.012$                             & $0.010 \pm 0.002$                             \\
                            &                               & 2                                & $0.830$                         & $-0.000$ $(p=0.552)$          & $0.016$                             & $0.017 \pm 0.002$                             \\
                            &                               & 3                                & $0.595$                         & $-0.015$ $(p=0.000)$          & $0.071$                             & $0.065 \pm 0.010$                             \\ \hline

@@

        \noindent Table \arabic{tablenum}: Estimates and standard errors for the parameters of the distribution of acceleration across all four models in the simulation study. The ``Estimate" column is the average across 500 simulations, and the ``Observed Fisher SE" column is the median across 500 simulations. The $\pm$ refers to the inter-quartile range. SE stands for standard error. Note that standard error cannot be estimated when $\phi^*_A \approx 0$ or $\phi^*_A \approx 1$. Reported $p$-values test if the observed bias is statistically significant using a one-sample $t$-test.
        \addtocounter{tablenum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of acceleration parameter estimates ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of acceleration parameter estimates ($\hat \mu^*_A$ and $\hat \sigma^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$ and $\sigma^*_A$ for both subdive states. These estimates are for the HHMM-DFT. Note that the HHMM-DFT assumes no auto-correlation in $\Zone_{t,\tilde t^*}$, so $\hat \phi_A^{*(\cdot,i^*)}$ is absent from these plots.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of acceleration parameter estimates ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHHMM. 
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distributions of acceleration parameter estimates ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. The red star represents the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for both subdive states. These estimates are for the CarHMM-DFT.
        \addtocounter{fignum}{1}

    \newpage
    \section{Empirical joint distributions - wiggliness ($\Ztwo_{t,\tilde t^*}$)}

@@

Model                       & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Subdive Type} & \multicolumn{1}{c}{Estimate}   & \multicolumn{1}{c}{Bias}   & \multicolumn{1}{c}{Empirical SE} & \multicolumn{1}{c}{Observed Fischer SE}       \\ \hline
\multirow{6}{*}{CarHHMM-DFT}& \multirow{3}{*}{$\mu_W^*$}    & 1                                & $34.981$                         & $0.001$ $(p=0.978)$          & $1.026$                             & $0.658 \pm 0.088$                             \\
                            &                               & 2                                & $505.10$                         & $-0.69$ $(p=0.394)$          & $17.94$                             & $9.74 \pm 0.81$                             \\
                            &                               & 3                                & $9787.1$                         & $17.2$ $(p=0.718)$           & $1060.5$                            & $262.1 \pm 97.120$                             \\
                            & \multirow{3}{*}{$\sigma_W^*$} & 1                                & $23.849$                         & $-0.001$ $(p=0.991)$         & $0.989$                             & $0.658 \pm 0.088$                             \\
                            &                               & 2                                & $514.21$                         & $-2.47$ $(p=0.010)$          & $21.43$                             & $9.74 \pm 0.81$                             \\
                            &                               & 3                                & $14430$                          & $-33$ $(p=0.647)$            & $1596$                              & $262 \pm 97$                             \\ \hline
\multirow{6}{*}{HHMM-DFT}   & \multirow{3}{*}{$\mu_W^*$}    & 1                                & $35.067$                         & $0.087$ $(p=0.097)$          & $1.170$                             & $0.679 \pm 0.100$                             \\
                            &                               & 2                                & $500.93$                         & $-4.86$ $(p=0.000)$          & $24.02$                             & $10.75 \pm 1.21$                             \\
                            &                               & 3                                & $6841.0$                         & $-2929.0$ $(p=0.000)$        & $1164.3$                             & $198.2 \pm 77.1$                             \\
                            & \multirow{3}{*}{$\sigma_W^*$} & 1                                & $23.800$                         & $-0.050$ $(p=0.351)$         & $1.192$                             & $0.679 \pm 0.100$                             \\
                            &                               & 2                                & $513.26$                         & $-3.42$ $(p=0.001)$          & $22.60$                             & $10.75 \pm 1.21$                             \\
                            &                               & 3                                & $11459$                          & $-3004$ $(p=0.000)$          & $1739$                              & $198 \pm 77$                             \\ \hline

@@

\multirow{6}{*}{CarHMM-DFT} & \multirow{3}{*}{$\mu_W^*$}    & 1                                & $35.034$                         & $0.054$ $(p=0.264)$          & $1.076$                             & $0.661 \pm 0.093$                             \\ 

>>>>

% Title, authors, affiliations
\title{Modelling multi-scale, state-switching functional data with hidden Markov models: Supplement B (simulation study)}
\date{}
\author{Evan Sidrow, Nancy Heckman, Sarah M. E. Fortune, \\ Andrew W. Trites, Ian Murphy, and Marie Auger-M\'eth\'e}

@@

    	\noindent Figure \arabic{fignum}: Acceleration data for five dives of a selected simulated data set. The raw accelerometer data is denoted by $Y^*_{t,\tilde t^*}$ and recorded at 50 $Hz$ while $\Zone_{t,\tilde t^*}$ and $\Ztwo_{t,\tilde t^*}$ are recorded at $0.5 Hz$. Dives are separated by vertical black lines. Curve colour corresponds to the true subdive state, while background colour corresponds to true dive type. Subdive state 1 can plausibly be interpreted as ``gliding", subdive state 2 can plausibly be interpreted as active swimming, and subdive state 3 can be interpreted as sudden or aggressive movement.
    	\addtocounter{fignum}{1}

    \newpage
    \section{Accuracy Comparisons}

@@

Model                       & \multicolumn{1}{c}{Training time ($min$)} & \multicolumn{1}{c}{Dive type} & \multicolumn{1}{c}{Subdive state} & \multicolumn{1}{c}{Dive accuracy} & \multicolumn{1}{c}{Subdive accuracy}  \\ \hline

@@

        \noindent Table \arabic{tablenum}: Average decoding accuracies and training times for all models used to categorize dive type and subdive state in the simulation study. Each of the four models was fit to 500 training data sets composed of 100 simulated dives and tested on test data sets also composed of 100 simulated dives. Reported values are averages, and figures following $\pm$ refer to sample standard deviation across the 500 data sets. ``All" denotes overall average decoding accuracy.
        \addtocounter{tablenum}{1}

    \newpage
    \section{Empirical Joint Distributions: Dive Duration ($Y_t$)}

@@

Model                       & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Dive type} & \multicolumn{1}{c}{Estimate}   & \multicolumn{1}{c}{Bias}   & \multicolumn{1}{c}{Empirical SE}   & \multicolumn{1}{c}{Observed Fisher SE}     \\ \hline
\multirow{4}{*}{CarHHMM-DFT}& \multirow{2}{*}{$\mu_Y$}      & 1                             & $27.33$                         & $0.07$ $(P=0.287)$          & $1.39$                             & $1.03 \pm 0.12$                             \\
                            &                               & 2                             & $127.14$                        & $0.07$ $(P=0.945)$          & $22.56$                             & $15.71 \pm 6.726$                             \\
                            & \multirow{2}{*}{$\sigma_Y$}   & 1                             & $10.80$                         & $-0.11$ $(P=0.035)$          & $1.15$                             & $0.84 \pm 0.11$                             \\
                            &                               & 2                             & $59.52$                         & $-4.54$ $(P=0.000)$          & $17.68$                             & $13.04 \pm 5.53$                             \\ \hline
\multirow{4}{*}{HHMM-DFT}   & \multirow{2}{*}{$\mu_Y$}      & 1                             & $27.37$                         & $0.10$ $(P=0.084)$          & $1.33$                             & $1.03 \pm 0.11$                             \\
                            &                               & 2                             & $126.84$                         & $-0.23$ $(P=0.818)$          & $22.00$                             & $15.37 \pm 6.11$                             \\
                            & \multirow{2}{*}{$\sigma_Y$}   & 1                             & $10.86$                         & $-0.05$ $(P=0.351)$          & $1.19$                             & $0.85 \pm 0.11$                             \\
                            &                               & 2                             & $57.14$                         & $-6.92$ $(P=0.000)$          & $16.12$                             & $12.62 \pm 5.17$                             \\ \hline
\multirow{4}{*}{CarHHMM}    & \multirow{2}{*}{$\mu_Y$}      & 1                             & $27.25$                         & $-0.02$ $(P=0.826)$          & $2.20$                             & $1.04 \pm 0.14$                             \\
                            &                               & 2                             & $117.62$                         & $-9.45$ $(P=0.000)$          & $34.34$                             & $15.13 \pm 7.10$                             \\
                            & \multirow{2}{*}{$\sigma_Y$}   & 1                             & $10.52$                         & $-0.39$ $(P=0.001)$          & $2.59$                             & $0.86 \pm 0.15$                             \\
                            &                               & 2                             & $56.98$                         & $-7.08$ $(P=0.000)$          & $16.33$                             & $12.53 \pm 5.95$                             \\ \hline

@@

        \noindent Table \arabic{tablenum}: Estimates and standard errors for the parameters of the distribution of dive duration across all four models in the simulation study. The ``Estimate" column gives averages across 500 simulations, and the ``Observed Fisher SE" column gives medians across 500 simulations. Figures following $\pm$ refer to inter quartile range. ``SE" stands for standard error. Reported $P$-values correspond to one-sample $t$-tests of whether bias is statistically significant.
        \addtocounter{tablenum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of the dive duration parameter estimators ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHHMM-DFT and correspond to both dive types. White stars represent the mean of all maximum likelihood estimates and red stars represent the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of the dive duration parameter estimators ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the HHMM-DFT and correspond to both dive types. White stars represent the mean of all maximum likelihood estimates and red stars represent the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of the dive duration parameter estimators ($\hat \mu_Y$ and $\hat \sigma_Y$). These estimates are associated with the CarHHMM and correspond to both dive types. White stars represent the mean of all maximum likelihood estimates and red stars represent the true values of $\mu_Y$ and $\sigma_Y$ for both dive types.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of the dive duration parameter estimators ($\hat \mu_Y$ and $\hat \sigma_Y$). This estimate is associated with the CarHMM-DFT. The White star represents the mean of all maximum likelihood estimates. Unlike in other figures, there is no red star representing true values because the CarHMM-DFT assumes that there is only one dive type.
        \addtocounter{fignum}{1}

    \newpage
    \section{Empirical Joint Distributions: Acceleration ($\Zone_{t,\tilde t^*}$)}

@@

Model                       & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Subdive state} & \multicolumn{1}{c}{Estimate}   & \multicolumn{1}{c}{Bias}   & \multicolumn{1}{c}{Empirical SE}   & \multicolumn{1}{c}{Observed Fisher SE}     \\ \hline
\multirow{9}{*}{CarHHMM-DFT}& \multirow{3}{*}{$\mu_A^*$}    & 1                                & $-0.012$                         & $-0.012$ $(P=0.004)$          & $0.093$                             & $0.066 \pm 0.038$                             \\
                            &                               & 2                                & $0.099$                         & $-0.001$ $(P=0.436)$          & $0.019$                             & $0.019 \pm 0.003$                             \\
                            &                               & 3                                & $0.200$                         & $0.000$ $(P=0.965)$          & $0.060$                             & $0.053 \pm 0.013$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.050$                         & $0.000$ $(P=0.376)$          & $0.002$                             & $0.002 \pm 0.000$                             \\
                            &                               & 2                                & $0.100$                         & $-0.000$ $(P=0.672)$          & $0.002$                             & $0.002 \pm 0.000$                             \\ 
                            &                               & 3                                & $0.298$                         & $-0.002$ $(P=0.024)$          & $0.016$                             & $0.015 \pm 0.002$                             \\
                            & \multirow{3}{*}{$\phi_A^*$}   & 1                                & $0.968$                         & $-0.002$ $(P=0.001)$          & $0.012$                             & $0.010 \pm 0.002$                             \\
                            &                               & 2                                & $0.828$                         & $-0.002$ $(P=0.007)$          & $0.017$                             & $0.017 \pm 0.002$                             \\
                            &                               & 3                                & $0.600$                         & $-0.010$ $(P=0.001)$          & $0.064$                             & $0.065 \pm 0.011$                             \\ \hline
\multirow{9}{*}{HHMM-DFT}   & \multirow{3}{*}{$\mu_A^*$}    & 1                                & $0.048$                         & $0.048$ $(P=0.000)$          & $0.033$                             & $0.007 \pm 0.001$                             \\
                            &                               & 2                                & $0.087$                         & $-0.013$ $(P=0.000)$          & $0.018$                             & $0.006 \pm 0.001$                             \\
                            &                               & 3                                & $0.215$                         & $0.015$ $(P=0.000)$          & $0.065$                             & $0.022 \pm 0.004$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.160$                         & $0.110$ $(P=0.000)$          & $0.023$                             & $0.005 \pm 0.001$                             \\
                            &                               & 2                                & $0.158$                         & $0.058$ $(P=0.000)$          & $0.010$                             & $0.004 \pm 0.000$                             \\ 
                            &                               & 3                                & $0.356$                         & $0.056$ $(P=0.000)$          & $0.035$                             & $0.015 \pm 0.002$                             \\

@@

\multirow{9}{*}{CarHHMM}    & \multirow{3}{*}{$\mu_A^*$}    & 1                                & $0.060$                         & $0.060$ $(P=0.000)$          & $0.034$                             & $0.019 \pm 0.006$                             \\
                            &                               & 2                                & $0.097$                         & $-0.003$ $(P=0.000)$          & $0.016$                             & $0.007 \pm 0.001$                             \\
                            &                               & 3                                & $0.201$                         & $0.001$ $(P=0.634)$          & $0.058$                             & $0.062 \pm 0.011$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.111$                         & $0.061$ $(P=0.000)$          & $0.004$                             & $0.003 \pm 0.000$                             \\
                            &                               & 2                                & $0.285$                         & $0.185$ $(P=0.000)$          & $0.009$                             & $0.005 \pm 0.000$                             \\ 
                            &                               & 3                                & $1.150$                         & $0.850$ $(P=0.000)$          & $0.080$                             & $0.048 \pm 0.009$                             \\
                            & \multirow{3}{*}{$\phi_A^*$}   & 1                                & $0.830$                         & $-0.140$ $(P=0.000)$          & $0.048$                             & $0.016 \pm 0.003$                             \\
                            &                               & 2                                & $0.004$                         & $-0.826$ $(P=0.000)$          & $0.012$                             & $0.001 \pm 0.023$                             \\
                            &                               & 3                                & $0.000$                         & $-0.610$ $(P=0.000)$          & $0.000$                             & $0.000 \pm 0.000$                             \\ \hline
\multirow{9}{*}{CarHMM-DFT} & \multirow{3}{*}{$\mu_A^*$}    & 1                                & $-0.008$                         & $-0.008$ $(P=0.142)$          & $0.114$                             & $0.064 \pm 0.037$                             \\
                            &                               & 2                                & $0.101$                         & $0.001$ $(P=0.438)$          & $0.019$                             & $0.019 \pm 0.003$                             \\
                            &                               & 3                                & $0.200$                         & $0.000$ $(P=0.926)$          & $0.057$                             & $0.054 \pm 0.014$                             \\
                            & \multirow{3}{*}{$\sigma_A^*$} & 1                                & $0.050$                         & $-0.000$ $(P=0.005)$          & $0.001$                             & $0.002 \pm 0.000$                             \\
                            &                               & 2                                & $0.100$                         & $-0.000$ $(P=0.000)$          & $0.002$                             & $0.002 \pm 0.000$                             \\ 
                            &                               & 3                                & $0.299$                         & $-0.001$ $(P=0.117)$          & $0.015$                             & $0.015 \pm 0.002$                             \\
                            & \multirow{3}{*}{$\phi_A^*$}   & 1                                & $0.968$                         & $-0.002$ $(P=0.001)$          & $0.012$                             & $0.010 \pm 0.002$                             \\
                            &                               & 2                                & $0.830$                         & $-0.000$ $(P=0.552)$          & $0.016$                             & $0.017 \pm 0.002$                             \\
                            &                               & 3                                & $0.595$                         & $-0.015$ $(P=0.000)$          & $0.071$                             & $0.065 \pm 0.010$                             \\ \hline

@@

        \noindent Table \arabic{tablenum}: Estimates and standard errors for the parameters of the distribution of acceleration across all four models in the simulation study. The ``Estimate" column gives averages across 500 simulations, and the ``Observed Fisher SE" column gives medians across 500 simulations. Figures following $\pm$ refer to inter quartile range. ``SE" stands for standard error. Note that standard error cannot be estimated when $\phi^*_A \approx 0$ or $\phi^*_A \approx 1$. Reported $P$-values correspond to one-sample $t$-tests of whether bias is statistically significant.
        \addtocounter{tablenum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimates of the distribution of acceleration parameter estimators ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. White stars represent the mean of all maximum likelihood estimates and red stars represent the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for each subdive state. These estimates are for the CarHHMM-DFT.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimate of the distribution of acceleration parameter estimators ($\hat \mu^*_A$ and $\hat \sigma^*_A$) for each subdive state. White stars represent the mean of all maximum likelihood estimates and red stars represent the true values of $\mu^*_A$ and $\sigma^*_A$ for each subdive state. These estimates are for the HHMM-DFT. Note that the HHMM-DFT assumes no autocorrelation in $\Zone_{t,\tilde t^*}$, so $\hat \phi_A^{*(\cdot,i^*)}$ is absent from these plots.
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimates of the distribution of acceleration parameter estimators ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. White stars represent the mean of all maximum likelihood estimates and red stars represent the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for each subdive state. These estimates are for the CarHHMM. 
        \addtocounter{fignum}{1}

@@

        \noindent Figure \arabic{fignum}: Kernel density estimates of the distribution of acceleration parameter estimators ($\hat \mu^*_A$, $\hat \sigma^*_A$, and $\hat \phi^*_A$) for each subdive state. White stars represent the mean of all maximum likelihood estimates and red stars represent the true values of $\mu^*_A$, $\sigma^*_A$, and $\phi^*_A$ for each subdive state. These estimates are for the CarHMM-DFT.
        \addtocounter{fignum}{1}

    \newpage
    \section{Empirical Joint Distributions: ``Wiggliness" ($\Ztwo_{t,\tilde t^*}$)}

        \begin{center}
        \scalebox{0.75}{
        \bgroup
        \centering
        \def\arraystretch{1.5}

@@

Model                       & \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Subdive state} & \multicolumn{1}{c}{Estimate}   & \multicolumn{1}{c}{Bias}   & \multicolumn{1}{c}{Empirical SE} & \multicolumn{1}{c}{Observed Fisher SE}       \\ \hline
\multirow{6}{*}{CarHHMM-DFT}& \multirow{3}{*}{$\mu_W^*$}    & 1                                & $34.981$                         & $0.001$ $(P=0.978)$          & $1.026$                             & $0.658 \pm 0.088$                             \\
                            &                               & 2                                & $505.104$                         & $-0.686$ $(P=0.394)$          & $17.936$                             & $9.740 \pm 0.814$                             \\
                            &                               & 3                                & $9787.146$                         & $17.166$ $(P=0.718)$          & $1060.450$                             & $262.144 \pm 97.120$                             \\
                            & \multirow{3}{*}{$\sigma_W^*$} & 1                                & $23.849$                         & $-0.001$ $(P=0.991)$          & $0.989$                             & $0.652 \pm 0.094$                             \\
                            &                               & 2                                & $514.207$                         & $-2.473$ $(P=0.010)$          & $21.427$                             & $11.410 \pm 0.947$                             \\
                            &                               & 3                                & $14429.732$                         & $-32.818$ $(P=0.647)$          & $1596.020$                             & $302.698 \pm 165.935$                             \\ \hline
\multirow{6}{*}{HHMM-DFT}   & \multirow{3}{*}{$\mu_W^*$}    & 1                                & $35.067$                         & $0.087$ $(P=0.097)$          & $1.170$                             & $0.679 \pm 0.100$                             \\
                            &                               & 2                                & $500.928$                         & $-4.862$ $(P=0.000)$          & $24.024$                             & $10.754 \pm 1.205$                             \\
                            &                               & 3                                & $6840.995$                         & $-2928.985$ $(P=0.000)$          & $1164.288$                             & $198.162 \pm 77.050$                             \\
                            & \multirow{3}{*}{$\sigma_W^*$} & 1                                & $23.800$                         & $-0.050$ $(P=0.351)$          & $1.192$                             & $0.676 \pm 0.104$                             \\
                            &                               & 2                                & $513.263$                         & $-3.417$ $(P=0.001)$          & $22.596$                             & $12.497 \pm 1.301$                             \\
                            &                               & 3                                & $11458.795$                         & $-3003.755$ $(P=0.000)$          & $1738.793$                             & $280.244 \pm 136.259$                             \\ \hline

@@

\multirow{6}{*}{CarHMM-DFT} & \multirow{3}{*}{$\mu_W^*$}    & 1                                & $35.034$                         & $0.054$ $(P=0.264)$          & $1.076$                             & $0.661 \pm 0.093$                             \\ 
                            &                               & 2                                & $503.762$                         & $-2.028$ $(P=0.013)$          & $18.258$                             & $9.641 \pm 0.916$                             \\
